### 2024-03-06 thesis/input_files/literaturereview.tex
6 march

@@

%SPIEGO UN PO' I PAPER CHE HO FATTO PASSARE VELOCEMENTE 
Every paper uses different datasets (heterogeneous)
So it is not possible to compare directly results
from one paper to another without implementing the
paper specific algorithms and the applying them to
your dataset.
This is why sections after are destined to analyzing
how these proposed methods so far work and their mathematical
theory details
\\
This problem has been already addressed in \cite{probablistic_electricity_forecast} and \cite{probablistic_electricity_forecast2}. 
These papers surveyed the performance of neural network architectures against simpler approaches like quantile regression and data fitting to Johnson distribution. Their conclusion is that distributional NN perform a little worse than quantile regression but the former has smaller computational cost; that is because the quantile regression is run for every quantile from 0.01 to 0.99.
Nevertheless kernel methods received very little attention in this specific setting.
\\
Kernel methods considered are kernel mean embedding \cite{pmlr}, \cite{Muandet_2017} and kernel herding \cite{supersamples}. Particularly extending the idea of \cite{2022nystrom}, where the Nyström approximation is employed in computing the kernel mean embedding, experiments with the Pivoted Cholesky decomposition will be performed.
\\
In section 2, three papers that lay the basis for this thesis’ work are summarized.
\\

>>>>

Energy Forecasting

\subsection{Energy forecasting classification}

\subsubsection{Types}
In the context of energy forecasting, the quantities of most interest are price, load and renewables generation.

\subsubsection{Forecasting horizons}
\subsubsection{Size}
- Forecasting horizons

- Refer to section \ref*{metrics} for evaluating metrics

- plots come quelli delle review con i dati di Scopus se é una cosa fattibile
\label{sec:literaturereview}

\begin{figure}[h!]
    \includegraphics[width=\textwidth]{images/epf_evolution1.jpg}
    \caption{EPF publications \cite{EPF_review}}
    \label{fig:epf_evolution}
  \end{figure}

\begin{figure}[h!]
  \includegraphics[width=\textwidth]{images/point_vs_prob.jpg}
  \caption{Point vs probabilistic pubblications}
  \label{fig:point_vs_prob}
\end{figure}

\begin{figure}[h!]
  \includegraphics[width=\textwidth]{images/cs_stat_both.jpg}
  \caption{Pubblications by method}
  \label{fig:point_vs_prob}
\end{figure}

\begin{figure}[h!]
  \includegraphics[width=\textwidth]{images/subject_area.jpg}
  \caption{Pubblications by subject area}
  \label{fig:subject_area}
\end{figure}

\begin{figure}[h!]
  \includegraphics[width=\textwidth]{images/src_title.jpg}
  \caption{Most popolar sources/outlets}
  \label{fig:src_title}
\end{figure}

@@

%SPIEGO UN PO' I PAPER CHE HO FATTO PASSARE VELOCEMENTE 
Every paper uses different datasets (heterogeneous)
So it is not possible to compare directly results
from one paper to another without implementing the
paper specific algorithms and the applying them to
your dataset.
This is why sections after are destined to analyzing
how these proposed methods so far work and their mathematical
theory details
\\
This problem has been already addressed in \cite{probablistic_electricity_forecast} and \cite{probablistic_electricity_forecast2}. 
These papers surveyed the performance of neural network architectures against simpler approaches like quantile regression and data fitting to Johnson distribution. Their conclusion is that distributional NN perform a little worse than quantile regression but the former has smaller computational cost; that is because the quantile regression is run for every quantile from 0.01 to 0.99.
Nevertheless kernel methods received very little attention in this specific setting.
\\
Kernel methods considered are kernel mean embedding \cite{pmlr}, \cite{Muandet_2017} and kernel herding \cite{supersamples}. Particularly extending the idea of \cite{2022nystrom}, where the Nyström approximation is employed in computing the kernel mean embedding, experiments with the Pivoted Cholesky decomposition will be performed.
\\
In section 2, three papers that lay the basis for this thesis’ work are summarized.
\\
************************
\\
During the past 25 years %because paper of 2014 says 15, so now it is 25
a wide range of new ideas have been proposed for point forecasting and for probabilistic forecasting.
\\
The field benefitted greatly from the increse of computing power, the greater availability of dataset and the interest in data science.
As a consequence, the forecaster's toolbox has grown in size and complexity.
\\
Such variety of methods is characterized by heterogeneity in the fields from which they come from; methods come from statisitics, mathematics, econometrics, electrical engineering and the artificial intelligence communities.
\\
Before delving into the literature review, it is important to make clear that at this point in time there is no superior method. Different solutions may outperform or underperform compared to other techniques depending on the problem settings. Thus, understanding the complexity, strenghts and weaknesses of each method is crucial for fitting the right model to the right setting.
\\
Finally, within this research community, emerged the need of more homogeneity in the choice of the error valuation metrics and in the way of comparing model performances \cite{EPF_review}.
\\
************************
\\
Lately, the idea of combining forecasts has gained popularity in the forecasting communit \cite{forecasting_big}; in the literature, combined forecasts are called ensemble \cite{gneiting_weather_ensemble}.
Experimental results have shown ensemble methods to outperform their component forecasts.
\\
Note that the more the errors of the combined models are not correlated the more we can benefit from ensembles.
\\
It is also worth noting that older and simpler methods are still valuable(in combination with other models or on their own); these being less subject to overfitting than complex models.

### 2024-03-06 thesis/input_files/problemdescription.tex
6 march

- Here explain in detail what this thesis is about.

@@

- why is it relevant.

- an outline of the thesis with the main points presented in each chapter.

>>>>

Individuals and organizations constantly face situations of uncertainty. Thus, the need of robust forecasting methods. Such methods are crucial to the process of taking informed decision and to strategic planning.
\\
The basic idea of forecasting is that we can extract knowledge from the past in order to make educated guess about the future. Consequently, the range of fields where forecasting can be applied is very wide.
In this thesis, our focus lies on applying forecasting to the energy sector. 
\\
\\
The reason of our decision to focus on the energy market is mainly motivated by the rapid changes it has experienced.
Over the last decades, electricity markets have gone through an unprecedented transformation; this shift was driven by the liberalization of such markets, the development and integration of renewable energy sources, the  increase of low carbon technologies and the adoption of smart meters. Events like the California electricity crisis help also motivating the choice of the electricity sector as subject of our studies, see \cite{california}.
Interestingly, the process of deregulation lead to an increasing interest in the field of electricity price forecasting (EF) within the academic community \ref{fig:epf_evolution}.
\\
Moreover, the United Nations have identified the right to access affordable, reliable, sustainable and modern energy as one of their 17 SDGs \cite{un_sdgs}
\\
Finally, the electricity market has a set of features that make it unique: electricity cannot be stored in an efficient way and supply and demand have to be matched instantly.
\\
\section{Motivation}
The are mutliple reasons why the energy sector needs robust forecasting techniques.
For power market companies, being able to predict prices with a low MAPE \ref{mape} results in increased savings \cite{savings}. Furthermore,the adoption of smart meters provides power market companies with a ton of consumer data; this will enable them to better model consumer preferences.
\\
Transmission system operators main goal is to match supply and deman, generally TSO do so by increasing or decreasing the generation supplied. Thus, from their view point forecasting is critical for balancing the electricity netwrok.
\\
Other possible applications are: control of storage, demand side response, anomaly detection, network design and planning, simulating inputs and hangling missing data.
\\
\section{Point vs probabilistic forecast}
A distinction has to be made between two types of forecasting approaches: point forecasts and probabilistic forecasts.
Point prediction, also called deterministic forecasting in the literature, is all about predicting a particular value in time.
On the other hand, with probabilistic forecasting we aim at predicting either a prediction interval, quantiles or a probability distribution for each point in time. For this reason, probabilistic forecasts are more informative than point forecasts; and this is why the interest of the research community is shifting towards them.
Note that a probabilistic forecast can be turned into a point forecast by simply taking its expecatation.
Alternatively, a probabilistic forecast can be derived from a point one by modeling the residuals of the point prediction.

\section{Aims and objectives}

@@

The scope of the thesis is analyzing state of the art forecasting methods in the energy market and to compare them with ideas coming from kernel theory. Applications of such methods to

\section{Outline}
We start with a literature review and bibliometric analysis in section \ref{literature_review}.
Then the theory underlying kernel methods is covered in section \ref{kernel_theory}. Following, section \ref{state_of_the_art} introduces the state of the art methods in the context of both point and probabilistic forecasting. Section \ref{energy_market} explains the core features and terminology of the energy market and of the electricity newtork.
Evaluation metrics necessary to rank the forecasting techniques are presented in section \ref{metrics}. Section \ref{eda} goes on with the ETL pipeline and the exploratory analysis. 
Implementation details are included in \ref{implementation}.
Finally section \ref{experiments} presents the experiments, the results and discusses models' strenghts, weaknesses and possible improvements.

### 2024-03-06 thesis/input_files/symbols.tex
6 march

\glsxtrnewsymbol[description={position}]{x}{\ensuremath{x}}
\glsxtrnewsymbol[description={velocity}]{v}{\ensuremath{v}}
\glsxtrnewsymbol[description={acceleration}]{a}{\ensuremath{a}}
\glsxtrnewsymbol[description={time}]{t}{\ensuremath{t}}
\glsxtrnewsymbol[description={force}]{F}{\ensuremath{F}}

>>>>

\glsxtrnewsymbol[description={acceleration}]{a}{\ensuremath{a}}

%C
\glsxtrnewsymbol[description={Continous ranked probability score}]{CRPS}{CRPS}

%E
\glsxtrnewsymbol[description={Exploratory data analysis}]{EDA}{EDA}
\glsxtrnewsymbol[description={Electric forecasting}]{EF}{EF}
\glsxtrnewsymbol[description={Electric vehicles}]{EVs}{EVs}
\glsxtrnewsymbol[description={Extract transform load}]{ETL}{ETL}

%L
\glsxtrnewsymbol[description={Low carbon technologies}]{LCTs}{LCTs}

\glsxtrnewsymbol[description={Low voltage}]{LV}{LV}

%P
\glsxtrnewsymbol[description={Prediction interval}]{PI}{PI}
\glsxtrnewsymbol[description={Price Forecasting}]{PF}{PF}

\glsxtrnewsymbol[description={Probabilistic Price Forecasting}]{PPF}{PPF}

%S
\glsxtrnewsymbol[description={Sustainable development goals}]{SDGs}{SDGs}

%
\glsxtrnewsymbol[description={Transmission system operator}]{TSO}{TSO}
